*Orchid fleck dichorhavirus*: a new *Brevipalpus* transmitted virus fresh from Florida

### 2021-03-29 ofv_presentation_2021.Rmd
updated

---
title: "*Orchid fleck dichorhavirus*: a new *Brevipalpus* transmitted virus fresh from Florida"
subtitle: "A. **Fife**^1^, D. **Carrillo**^2^, G. **Knox**^3^, F. **Iriarte**^4^, K. **Dey**^5^, A. **Roy**^6^, R. **Ochoa**^7^, G. **Bauchan**^8^, M. **Paret**^4,9^, X. **Martini**^1^*"
author: "^1^ University of Florida, Department of Entomology and Nematology, North Florida Research and Education Center, Quincy FL 32351 ^2^ University of Florida, Department of Entomology and Nematology, Tropical Research and Education Center, Homestead FL 33031 ^3^ University of Florida, Department of Environmental Horticulture, North Florida Research and Education Center, Quincy FL 32351 ^4^ University of Florida, Department of Plant Pathology, North Florida Research and Education Center, Quincy FL 32351 ^5^ The Florida Department of Agriculture and Consumer Services, Division of Plant Industry, Section of Plant Pathology, Doyle Conner Building, 1911 SW 34th street, Gainesville, FL 32608 ^6^	United States Department of Agriculture – Agriculture Research Service, Molecular Plant Pathology Laboratory, 10300 Baltimore Ave, Bldg. 4 BARC-West, Beltsville, MD 20705 ^7^ United States Department of Agriculture - Agriculture Research Service, Systematic Entomology Laboratory 10300 Baltimore Ave, Bldg. 5 BARC-West, Beltsville, MD 20705 ^8^	United States Department of Agriculture - Animal and Plant Health Inspection Service, Electron and Confocal Microscopy Unit, Bldg. 12  BARC-West, 10300 Baltimore Ave, Beltsville, MD 20705 ^9^ Plant Pathology Department, University of Florida, Gainesville, FL 32611"
date: March 29, 2021
output: 
  powerpoint_presentation
bibliography: 'thesis.bib'
csl: 'j_med_entomol.csl'
#always_allow_html: true
---
```{r setup, include=FALSE}
#setup options
knitr::opts_chunk$set(fig.width=12, fig.height=8, echo=FALSE)

#a list of packages used for scripts
pkgs <- c('captioner', 'pryr', 'tiff')

@@

# What was causing these symptoms?
:::::::::::::: {.columns}
::: {.column}
 - Tested negative for begomovirus, potyvirus, tospovirus as well as for Impatiens necrotic spot virus, Tobacco mosaic virus and Tomato spotted wilt virus
 - Sent to the Florida Department of Agriculture and Consumer Services (FDACS) $\rightarrow$ sent to USDA
:::
::: {.column}
![Symptoms on varigated Ophiopogonae](images/symp-ofv-lir-2.jpg)
:::
::::::::::::::

@@

# Distribution of OFV
:::::::::::::: {.columns}
::: {.column}
content
:::
::: {.column}
 - First described infecting *Cymbidium* orchids in Japan [@Doi1977]
 - OFV and OFV-like rhabdoviruses have been reported worldwide
 - Asia, Africa, North America, South America, Europe and Oceania
 - Prevalence of OFV and its mite vector is thought to be associated with the movement of infected orchids [@Dietzgen2018a]
:::
::::::::::::::

# Maps
:::::::::::::: {.columns}
::: {.column}
 - Additional samples were taken from other Nolinoidaea, including *Liriope muscari*, *Ophiopogon japonicus*, *O. intermedius* and *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) in Leon and Alachua counties.
:::
::: {.column}
content
:::
::::::::::::::

# Mites Recovered
:::::::::::::: {.columns}
::: {.column}
 - Three *Brevipalpus* mite species were recovered from OFV-infected plants: 
  - *Brevipalpus californicus* sensu lato
  - *B. obovatus* Donnadieu 
  - *B. confusus* Banks
- One of these species is presumably responsible for OFV transmission
  - *Brevipalpus californicus* (Banks) group are the only known to transmit OFV in a persistent propagative manner
  - *Brevipalpus* mite species complex contains cryptic species [@Childers2011]
:::
::: {.column}
content
:::
::::::::::::::

# OFV is an issue for Florida agriculture
:::::::::::::: {.columns}
::: {.column}
 - The dichorhavirus that infects citrus in Hawaii, Mexico, Colombia, and South Africa are identical to the OFV in gene order, content, and the genome sequence.
 - ICTV: Officially, OFV consists of two orchid strains (OFV-Orc1 and OFV-Orc2) and two citrus strains (OFV-Cit1 and OFV-Cit2)
 - Orchid strains of OFV infects citrus [@Roy2020], but none of the citrus strains have been reported from any orchid species
 - OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping
:::
::: {.column}
content
:::
::::::::::::::

# Further considerations
 - Florida is also home to a plethora of native and naturalized orchid species
  - Cultivating *Vanilla* in southern Florida [@Chambers2019] 
  - Famous Ghost Orchid, [*Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe].
 - Citrus leprosis was present in Florida during the 1860’s and eradicated in the mid-1960s
 - @Kitajima2011 found that the Citrus Leprosis virus (CiLV) which previously affected Florida citrus was a nuclear type of citrus leprosis closely related to OFV strains [@Roy2013]
  - Association of a distant relative of OFV named Citrus leprosis dichorhavirus-N0 (CiLV-N0) was confirmed in relation to the leprosis disease outbreak in Florida [@Hartung2015].

# Threat for the multi-billion dollar citrus industry
  - Detection of OFV-Orc1 in South Africa [@Cook2019] in *C. sinensis* (Navel and Valencia orange) and OFV-Orc2 in Hawaii [@Velarde2021] in *C. reticulata* (mandarin) and *C. jambhiri* (rough lemon) highlights the threat of different strains of OFV
  - Lastly, some OFV isolates are known to be involved with citrus leprosis disease in Mexico [@Roy2015]

# Mites: a large part of the problem
:::::::::::::: {.columns}
::: {.column}
- Florida has various mite species of *Brevipalpus*:
- *B. californicus* and *B. yothersi* are both known vectors of Dichorhaviruses (OFV) and Cileviruses (Citrus Leprosis) [@Knorr1968; @Kondo2003; @Beltran-Beltran2020]
- *B. obovatus* is a suspected vector as well [@Childers2003]
- *Brevipalpus* mites are known to have cryptic species complexes and unresolved questions about host specificity
- Critical to identify the vector of OFVs in Florida and monitor its spread to determine the risk this virus represents for the native plants, agriculture and the ornamental/landscaping industries of Florida and the surrounding regions.
:::
::: {.column}
content
:::
::::::::::::::

# Conclusions
 - Three new hosts from the family Asperagaceae
 - Need to know the extended host range of OFVs
  - Survey of Florida's citrus growing regions, esp. plants within the families Rutaceae and Asparagaceae
 - Mite species determination: PCR methods?
 - Mite transmission assays

>>>>

---
title: "*Orchid fleck dichorhavirus*: a new *Brevipalpus* transmitted virus fresh from Florida"
subtitle: "A. **Fife**^1^, D. **Carrillo**^2^, G. **Knox**^3^, F. **Iriarte**^4^, K. **Dey**^5^, A. **Roy**^6^, R. **Ochoa**^7^, G. **Bauchan**^8^, M. **Paret**^4,9^, X. **Martini**^1^*"
author: "^1^ University of Florida, Department of Entomology and Nematology, North Florida Research and Education Center, Quincy FL 32351 ^2^ University of Florida, Department of Entomology and Nematology, Tropical Research and Education Center, Homestead FL 33031 ^3^ University of Florida, Department of Environmental Horticulture, North Florida Research and Education Center, Quincy FL 32351 ^4^ University of Florida, Department of Plant Pathology, North Florida Research and Education Center, Quincy FL 32351 ^5^ The Florida Department of Agriculture and Consumer Services, Division of Plant Industry, Section of Plant Pathology, Doyle Conner Building, 1911 SW 34th street, Gainesville, FL 32608 ^6^	United States Department of Agriculture – Agriculture Research Service, Molecular Plant Pathology Laboratory, 10300 Baltimore Ave, Bldg. 4 BARC-West, Beltsville, MD 20705 ^7^ United States Department of Agriculture - Agriculture Research Service, Systematic Entomology Laboratory 10300 Baltimore Ave, Bldg. 5 BARC-West, Beltsville, MD 20705 ^8^	United States Department of Agriculture - Animal and Plant Health Inspection Service, Electron and Confocal Microscopy Unit, Bldg. 12  BARC-West, 10300 Baltimore Ave, Beltsville, MD 20705 ^9^ Plant Pathology Department, University of Florida, Gainesville, FL 32611"
date: March 29, 2021
output: 
  powerpoint_presentation
bibliography: 'thesis.bib'
csl: 'j_med_entomol.csl'
always_allow_html: true
---
```{r setup, include=FALSE}
#setup options
knitr::opts_chunk$set(fig.width=12, fig.height=8, echo=FALSE)

#a list of packages used for scripts
pkgs <-
  c(
    'captioner',
    'pryr',
    'tiff',
    'sf',
    'rgeos',
    'ggspatial',
    'rnaturalearth',
    'rnaturalearthdata',
    'tidyverse',
    'ggmap',
    'ggplot2',
    'ggthemes',
    'extrafont',
    'showtext',
    'leaflet',
    'dplyr',
    'readxl',
    'maps',
    'mapview',
    'htmlwidgets',
    'viridis',
    'RColorBrewer'
  )

@@

# What was causing these symptoms?
:::::::::::::: {.columns}
::: {.column}
 - Tested negative for begomovirus, potyvirus, tospovirus as well as for *Impatiens necrotic spot virus*, *Tobacco mosaic virus* and *Tomato spotted wilt virus*
 - Sent to the Florida Department of Agriculture and Consumer Services (FDACS) $\rightarrow$ sent to USDA
:::
::: {.column}
![Symptoms on varigated Ophiopogonae](images/symp-ofv-lir-2.jpg)
:::
::::::::::::::

@@

# Worldwide Distribution of OFV
 - First described infecting *Cymbidium* orchids in Japan [@Doi1977]
 - OFV and OFV-like rhabdoviruses have been reported worldwide
 - Asia, Africa, North America, South America, Europe and Oceania
 - Prevalence of OFV and its mite vector is thought to be associated with the movement of infected orchids [@Dietzgen2018a]

# World Map of OFV Distribution
```{r, warning=FALSE, echo=FALSE, comment=FALSE}
#https://slcladal.github.io/maps.html
library(maptools)
data(wrld_simpl)
myCountries = wrld_simpl@data$NAME %in% c("China", "Korea, Republic of", "Korea, Democratic People's Republic of", "South Africa", "United States", "Mexico", "Brazil", "Colombia", "Costa Rica", "Paraguay", "Denmark", "France", "Germany", "Australia", "Fiji", "Vanuatu", "Japan")
plot(wrld_simpl, col = c(gray(.90), "orange")[myCountries+1])
```

# Map of OFV distribution in Tallahassee 
```{r}
#####LOADING REQUIRED FILES####
df <- read_xlsx('data/ofv_clean_datasheet.xlsx')

#making sure R knows what variables are factors
df$symptoms <- as_factor(df$symptoms)
df$plant_cv <- as_factor(df$plant_cv)
df$family <- as_factor(df$family)
df$order <- as_factor(df$order)

####REMOVING NAS####
#removing columns of NA
df <- drop_na(df)

#makes the dataframe into a spatial dataframe
df <-
  st_as_sf(df,
           coords = c("lon", "lat"),
           crs = 4326)

####GRAPHING SAMPLING LOCATIONS####
#creating a subset of the data containing only asparagales
asp <- df %>%
  filter(order == 'asparagales')

#making color pallet for the map
#the order goes (ringspot, stippling, none, chlorosis)
pal <- c('#092C48', '#A3B7F9', '#EDE6DE', '#C59434')

#mapping locations
mapview(
  asp,
  zcol = 'symptoms',
  col.regions = pal,
  legend = TRUE,
  alpha.regions = 1,
  layer.name = 'Ringspot Symptoms on Asparagales',
  map.types = 'Stamen.TonerLite'
)
```

# Mites Recovered During Our Survey
:::::::::::::: {.columns}
::: {.column}
- Unidentified eriophyid mite
- Three *Brevipalpus* mite species were recovered from OFV-infected plants: 
  - *Brevipalpus californicus* sensu lato
  - *B. obovatus* Donnadieu 
  - *B. confusus* Banks
- One of these species is presumably responsible for OFV transmission
- **Cryo-SEM to help determine *Brevipalpus* species identity**
:::
::: {.column}
![Various *Brevipalpus* spp. live in Florida](images/other_mites.tif)
:::
::::::::::::::

# The Main Suspect: *B. californicus* sensu lato
![*Brevipalpus californicus* s.l](images/b_cali_detail.tif)
- *Brevipalpus californicus* (Banks) group are the only known to transmit OFV in a persistent propagative manner
  - *Brevipalpus* mite species complex contains cryptic species $\rightarrow$ develop molecular methods?

# OFV is an issue for Florida agriculture
:::::::::::::: {.columns}
::: {.column}
 - The dichorhavirus that infects citrus in Hawaii, Mexico, Colombia, and South Africa are identical to the OFV in gene order, content, and the genome sequence.
 - ICTV: Officially, OFV consists of two orchid strains (OFV-Orc1 and OFV-Orc2) and two citrus strains (OFV-Cit1 and OFV-Cit2)
 - Orchid strains of OFV infects citrus [@Roy2020], but none of the citrus strains have been reported from any orchid species
 - OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping
:::
::: {.column}
```{r}
grid::grid.raster(tiff::readTIFF("images/NPA-20-02-Orchid-fleck-virus_Page_1.tiff"))
```
:::
::::::::::::::
